### 2015-07-02 Projects/DelCo_2012/Report/DelCo_2012.tex
Delco_2012:  update report

\begin{table}[!ht]
\centering
\captionof{table}{Operating conditions for experiments}\label{tab:op_condition}
\begin{tabular}{ll}
\toprule[1.5pt]
Equipment    &   Operating Condition \\
\midrule
Nozzle       & Metro 1, fixed gallonage fog,straight, or 7/8~inch solid stream \\
Hoseline     & 100~ft of 1-3/4~inch hoseline \\
Foam         & 0.3\% foam concentration \\
Pressure     & Water and Air pressure of 100 - 120~psi \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

Twenty-one of the tests were conducted in the concrete burn building where the gas cooling tests would take place, and eleven of the tests were conducted in the experimental structures to be used for the fire suppression tests. The test variables were the nozzle height, nozzle location within the test structure, and the nozzle pattern (cf. Fig.~\ref{fig:Spray_Density_Nozzle_Patterns}). A description and drawings of these facilities are included in Section~\ref{sec:Experimental Structures}. Detailed discussion of the tests and results are in Section~\ref{sec:Spray_Density}.

@@

Wood pallets were used to provide the fuel for the fire used in the gas cooling experiments.  The pallets were approximately   1.2 m (4.0 ft) by 1.0 m (3.3 ft) by 0.13 m (0.42 ft) thick.  Forty eight of the pallets used to fuel the gas cooling experiments were weighed.  The pallets ranged in mass from 13.6 kg (29.9 lbs) to 26.4 kg (58.1 lbs) with an average of 18.4 kg (40.5 lbs).  The initial fuel load consisted of 10 pallets, arranged in two stacks of five as shown in Fig xx.  Approximately one half of a bale of excelsior, 13.0  kg (28.6 lbs), was mixed with the pallets to aid with ignition.

As the pallets burned burn away and the hot gas layer temperatures decreased, the steel shutters on the window to the fire room were opened and additional pallets were added to the piles of burning pallets until the flames from the piles reached the ceiling of the burn room again.  The number of pallets added each time varied.

@@

\subsection{Fire Suppression}
\label{sec:Fuel_Load_Wood_Experiments}

The fuel load for the experiments consisted of one bale of hay, twelve pallets, eleven and a half sheets plywood on the walls and the wood ``floor assembly". Two stacks of pallets with hay were used as the first items ignited in each experiment.  Each stack was composed of six pallets with a half bale of hay.  The hay was layered between each pallet.  The pallets and hay were weighed prior to each experiment.  For the four experiments, the average mass of the two stacks of pallets and the bale of hay was 232.4~kg (511.3~lb).  A complete listing of the mass of each pallet and hay bales is given in Appendix ????????

The pallet stacks were located 0.3~m (12~in) away from the east and south walls of the burn room.  The pallets were 1.2~m (48~in) long, 1.0~m (40~in) and 123~mm (4.8~in) high.  The stacks were 150~mm (6~in) apart.  One layer of 13 mm (0.5 in) thick gypsum board panels were laid on the concrete floor under the wood pallets to form a protective layer to minimize thermal damage to the concrete floor.
The east and south wall of the burn room was covered with 15~mm (0.59~in) thick, sheets of plywood.  The ``wood flooring assembly", composed of 12 TGIs and 9 sheets of OSB, served as the ceiling of the burn room.  The open doorway to the burn room on the south side was covered with a 2.4~m (8~ft) tall by 1.2~m (4~ft) wide piece of medium density fiberboard paneling that was 5~mm (0.19~in) thick.  The southeast corner of the burn room with the fuel load installed is shown in Fig.~\ref{fig:Wood_Fuel_Load}.

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Pictures/Wood_Fuel_Package}
	\caption{Photograph of Southeast corner of burn room with wood fuel load.}
	\label{fig:Wood_Fuel_Load}
\end{figure}

@@

\begin{sidewaystable}[!ht]
	\centering
	\scriptsize
	\caption{Fire Suppression Fuel masses}
	\renewcommand{\tabcolsep}{1pt}
	\begin{tabular}{lllllll}
		\hline\noalign{\smallskip}
		Item               & Component		& Quantity		&  Material Description             							&  Dimensions (m)            	&  Mass (kg)  		& Total Mass (kg) \\
		\noalign{\smallskip}\hline\noalign{\smallskip}
		2-seat New Sofa    &				& 2				& 85~\% Urethane Foam, 15~\% Polyester Fiber and Wood Frame 	&  1.68 L X 0.90 D X 0.86 H  	&  72.4    			& 144.8 \\
		           	 	   & \multirow{2}{*}{{Mattress}}	    & \multirow{2}{*}{{1 per sofa}}	& \multirow{2}{*}{\parbox{5cm} {Cover: 52~\% Polypropylene, 48~\% Polyester, Filling: 66~\% Polyester Fiber Pad, 34~\% Polyester Fiber  Batting}} & \multirow{2}{*}{{1.83 L X 1.21 W X 0.14 H}}  & \multirow{2}{*}{{14.6}}     \\
		           	 	   &	            &               &                                                          		&                            	&             		& \\[.25cm]
			           	   & Seat Cushion	& 2 per sofa	& 85~\% Urethane Foam, 15~\% Polyester Fiber   				    &  0.69 L X 0.58 D X 0.14 H  	&  2.2     			& \\
		Blue Chair         &				& 2				& 85~\% Urethane Foam, 15~\% Polyester Fiber and Wood Frame   	&  0.81 L X 0.75 D X 0.91 H  	&  17.4    			& 34.8 \\
		    			   & Seat Cushion   & 1 per chair   & 85~\% Urethane Foam, 15~\% Polyester Fiber					&  0.51 L X 0.57 D X 0.14 H  	&  1.7    			& \\
		Blue/Green Chair   & 				& 2				& 85~\% Urethane Foam, 15~\% Polyester Fiber and Wood Frame     &  0.77 L X 0.71 D X 0.89 H  	&  16    			& 32 \\
		                   & Seat Cushion	& 1 per chair	& 85~\% Urethane Foam, 15~\% Polyester Fiber		            &  0.55 L X 0.53 D X 0.15 H  	&  1.6    			& \\
		Paneling           &				& 8 panels		& 3.2 mm thick medium density fiberboard                   		&  2.44 L X 1.22 W X 3.2mm thick&  9    			& 72 \\
		Carpeting          &				& 22.26 m$^2$	& 100~\% PET Polyester                  						&  3.68 L X 6.05 W  			&  1.44 kg/m$^2$	& 32.2 \\
		Carpet Padding     &				& 22.26 m$^2$	& Polyurethane                 	 								&  3.68 L X 6.05 W X 11mm thick &  0.89 kg/m$^2$    & 19.8 \\
		                   &                &               &                                                               &                               & Total Mass        & 335.6 \\
		\noalign{\smallskip}\hline
	\end{tabular}
	\label{tab:Fire_Suppression_Fuel_Masses}
\end{sidewaystable}

@@

The experiments were conducted by placing 2.5 ft$^2$ interlocking collector bins across the floor surface area of the burn rooms within each structure. The building was sheathed in plastic for the spray density tests to protect and preserve as much of the structure as possible before any fire suppression tests were conducted. Water flowed for a specified duration and upon completion the bins were removed from the structure. Each bin was then weighed to quantify the impact of nozzle location and spray pattern.

12 Experiments:
4 SS from room
2 Fog from room
2 SS from Hall
2 Fog from Hall
1 SB from Hall
1 SB from Room

Spray Distribution Summary:
- Experiments provided nozzle placement for gas cooling and fire suppression experiments.
- Provide a sense of the distribution of the water throughout the test space.

Hallway nozzle = gas cooling

Room nozzle = seat of fire

@@

>>>>

\begin{table}[!ht]
\centering
\captionof{table}{Operating conditions for experiments}\label{tab:op_condition}
\begin{tabular}{ll}
\toprule[1.5pt]
Equipment    &   Operating Condition \\
\midrule
Nozzle       & Metro 1, fixed gallonage fog, straight, or 7/8~inch solid stream \\
Hoseline     & 100~ft of 1-3/4~inch hoseline \\
Foam         & 0.3\% foam concentration \\
Pressure     & Water and Air pressure of 100 - 120~psi \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

Twenty-one of the tests were conducted in the concrete burn building where the gas cooling tests would take place, and twelve of the tests were conducted in the experimental structures to be used for the fire suppression tests. The test variables were the nozzle height, nozzle location within the test structure, and the nozzle pattern (cf. Fig.~\ref{fig:Spray_Density_Nozzle_Patterns}). A description and drawings of these facilities are included in Section~\ref{sec:Experimental Structures}. Detailed discussion of the tests and results are in Section~\ref{sec:Spray_Density}.

@@

Wood pallets were used as the fuel source for the gas cooling experiments. The pallets were approximately 1.2~m (4.0~ft) by 1.0~m (3.3~ft) by 0.13~m (0.42~ft) thick. The pallets ranged in mass from 13.6~kg (29.9~lb) to 26.4~kg (58.1~lb) with an average of 18.4~kg (40.5~lbs). The initial fuel load consisted of 10 pallets, arranged in two stacks of five as shown in Fig.~\ref{fig:Pallet_Fuel_Load}. Approximately one half of a bale of excelsior, 13.0~kg (28.6~lb), was mixed with the pallets to aid with ignition. Fig.~\ref{fig:Burn_Building_Fuel_Load}, provides dimensioned locations of the two stacks of pallets within the burn room.

@@

Note, that as the pallets burned away and the hot gas layer temperatures decreased, the steel shutters on the window to the burn room were opened and additional pallets were added. Pallets were added to the existing fuel locations until the flames from the piles reached the ceiling of the burn room again. The number of pallets added each time varied.

\subsection{Fire Suppression}
\label{sec:Fuel_Load_Fire_Suppression}

\subsubsection{Wood Fuel Package}
\label{sec:fire_suppression_pallet_fuel}

The fuel load for the experiments consisted of one bale of hay, twelve pallets, eleven and a half sheets plywood on the walls and the wood ``floor assembly". Two stacks of pallets with hay were used as the first items ignited in each experiment. Each stack was composed of six pallets with a half bale of hay. The hay was layered between each pallet. Figure~\ref{fig:Wood_Fuel_Load} shows the pallets prior to ignition. The pallets and hay were weighed prior to each experiment. For the four pallet experiments, the average mass of the two stacks of pallets and the bale of hay was 232.4~kg (511.3~lb). A complete listing of the mass of each pallet and hay bales is given in Appendix~\ref{app:fuel_loads}.

\begin{figure}[!ht]
	\includegraphics[width=0.65\columnwidth]{../Figures/Pictures/Wood_Fuel_Package}
	\caption{Photograph of Southeast corner of burn room with wood fuel load.}
	\label{fig:Wood_Fuel_Load}
\end{figure}

The pallet stacks were located 0.3~m (12~in) away from the east and south walls of the burn room (cf. Fig.~\ref{fig:Wood_Fuel_Load_Dimensions}). The pallets were 1.2~m (48~in) long, 1.0~m (40~in) and 123~mm (4.8~in) high. The stacks were spaced 150~mm (6~in) apart. One layer of 13~mm (0.5~in) thick gypsum board panels were laid on the concrete floor under the wood pallets to form a protective layer to minimize thermal damage to the concrete floor. The east and south wall of the burn room was covered with 15~mm (0.59~in) thick, sheets of plywood.  The ``wood flooring assembly", composed of 12 TGIs and 9 sheets of OSB, served as the ceiling of the burn room. The open doorway to the burn room on the south side was covered with a 2.4~m (8~ft) tall by 1.2~m (4~ft) wide piece of medium density fiberboard paneling that was 5~mm (0.19~in) thick.

@@

\subsubsection{Furniture Fuel Package}
\label{sec:fire_suppression_furniture_fuel}

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Pictures/Furniture_Fuel_Load}
	\caption{Furniture Fuel Load Dimensions.}
	\label{fig:Furniture_Fuel_Load}
\end{figure}

@@

\begin{sidewaystable}[!ht]
	\centering
	\scriptsize
	\caption{Fire Suppression Fuel masses}
	\renewcommand{\tabcolsep}{1pt}
	\begin{tabular}{lllllcc}
		\toprule[1.5pt]
		Item               & Component		& Quantity		&  Material Description             							&  Dimensions (m)            	&  Mass (kg)  		& Total Mass (kg) \\
		\midrule
		2-seat New Sofa    &				& 2				& 85~\% Urethane Foam, 15~\% Polyester Fiber, Wood Frame 	   &  1.68 L X 0.90 D X 0.86 H  	&  72.4    			& 144.8 \\
		           	 	   & \multirow{2}{*}{{Mattress}}	    & \multirow{2}{*}{{1 per sofa}}	& \multirow{2}{*}{\parbox{5cm} {Cover: 52~\% Polypropylene, 48~\% Polyester, Filling: 66~\% Polyester Fiber Pad, 34~\% Polyester Fiber  Batting}} & \multirow{2}{*}{{1.83 L X 1.21 W X 0.14 H}}  & \multirow{2}{*}{{14.6}}     \\
		           	 	   &	            &               &                                                          		&                            	&             		& \\[.25cm]
			           	   & Seat Cushion	& 2 per sofa	& 85~\% Urethane Foam, 15~\% Polyester Fiber   				    &  0.69 L X 0.58 D X 0.14 H  	&  2.2     			& \\
		Blue Chair         &				& 2				& 85~\% Urethane Foam, 15~\% Polyester Fiber, Wood Frame   	    &  0.81 L X 0.75 D X 0.91 H  	&  17.4    			& 34.8 \\
		    			   & Seat Cushion   & 1 per chair   & 85~\% Urethane Foam, 15~\% Polyester Fiber					&  0.51 L X 0.57 D X 0.14 H  	&  1.7    			& \\
		Blue/Green Chair   & 				& 2				& 85~\% Urethane Foam, 15~\% Polyester Fiber, Wood Frame        &  0.77 L X 0.71 D X 0.89 H  	&  16    			& 32 \\
		                   & Seat Cushion	& 1 per chair	& 85~\% Urethane Foam, 15~\% Polyester Fiber		            &  0.55 L X 0.53 D X 0.15 H  	&  1.6    			& \\
		Paneling           &				& 8 panels		& 3.2 mm thick medium density fiberboard                   		&  2.44 L X 1.22 W X 3.2~mm H   &  9    			& 72 \\
		Carpeting          &				& 22.26 m$^2$	& 100~\% PET Polyester                  						&  3.68 L X 6.05 W  			&  1.44 kg/m$^2$	& 32.2 \\
		Carpet Padding     &				& 22.26 m$^2$	& Polyurethane                 	 								&  3.68 L X 6.05 W X 11~mm H    &  0.89 kg/m$^2$    & 19.8 \\
		                   &                &               &                                                               &                               & Total Mass        & 335.6 \\
		\bottomrule[1.25pt]
	\end{tabular}
	\label{tab:Fire_Suppression_Fuel_Masses}
\end{sidewaystable}

@@

Spray density experiments were conducted by placing 2.5 ft$^2$ interlocking collector bins across the floor surface area of the burn rooms within each structure. The buildings were sheathed in plastic for the spray density tests to protect and preserve as much of the structure as possible before any fire suppression tests were conducted. Water flowed for a specified duration and upon completion the bins were removed from the structure. Each bin was then weighed to quantify the impact of nozzle location and spray pattern. Thirty-three spray density tests were conducted; 21 were in the burn building and 12 were in the concrete structure. Table~\ref{tab:spray_density_tests} provides and overview of each individual experiment. In the table, BB stands for the burn building tests while CS stands for the concrete structure tests. For all of the experiments, 100~ft (30~m) of 1~3/4~inch hoseline connected to a Blitz Fire monitor with a Metro 1 nozzle was used.

\begin{table}[!ht]
\footnotesize
\centering
\captionof{table}{Spray Density Test Matrix}\label{tab:spray_density_tests}
\begin{tabular}{lllllll}
\toprule[1.5pt]
Test \#    &   Flow Rate (gpm)  & Duration (s)  & Fluid  &  Pattern            & Nozzle Height (in) & Location (in) \\
\midrule
BB1        &   120              & 15            & Water  &  Solid Stream back  & 20~7/8             &          \\
BB2        &   120              & 23            & Water  &  Solid Stream       &                    &          \\
BB3        &   120              &               & Water  &  Solid Stream       & 16                 &          \\
BB4        &   120              &               & Water  &  Solid Stream       & 16                 &          \\
BB5        &   120              & 20            & Water  &  Solid Stream       & 16                 &          \\
BB6        &   120              & 21            & Water  &  30$^{\circ}$ fog   & 15                 &          \\
BB7        &   120              & 21            & Water  &  30$^{\circ}$ fog   & 15                 &          \\
BB8        &   120              & 22            & Water  &  30$^{\circ}$ fog   & 17                 &          \\
BB9        &   120              & 24            & Water  &  30$^{\circ}$ fog   & 17~1/2             &          \\
BB10       &   120              & 20            & Water  &  Solid 7/8~in slug  & 12~1/2             &          \\
BB11       &   120              & 22            & Water  &  Solid 7/8~in slug  & 11~1/2             &          \\
BB12       &   120              & 15            & CAF    &  Smooth Bore 7/8~in & 12~1/2             &          \\
BB13       &   120              & 10            & CAF    &  Smooth Bore 7/8~in &                    &          \\
BB14       &   120              & 15            & CAF    &  Smooth Bore 7/8~in &                    &          \\
BB15       &   120              & 20            & CAF    &  Smooth Bore 7/8~in &                    &          \\
BB16       &   120              &               & CAF    &  30$^{\circ}$ fog   & 15                 &          \\
BB17       &   120              & 20            & CAF    &  Solid Stream narrow & 15                &          \\
BB18       &   120              & 10            & CAF    &  Solid Stream mid   &                    &          \\
BB19       &   120              & 15            & CAF    &  30$^{\circ}$ fog   & 15                 &          \\
BB20       &   120              &               & CAF    &  30$^{\circ}$ fog   & 17                 &          \\
BB21       &   120              &               & CAF    &  Solid Stream back  &                    &          \\
CS1        &   115,75           &               & Water  &  Solid Stream       &                    &          \\
CS2        &   120,115          & 18            & Water  &  Solid Stream       &                    &          \\
CS3        &   120              & 19            & Water  &  Solid Stream       &                    &          \\
CS4        &   120,118          & 21            & Water  &  Solid Stream       &                    &          \\
CS5        &   120,118          & 22            & Water  &  30$^{\circ}$ fog   &                    &          \\
CS6        &   120,118          & 19            & Water  &  30$^{\circ}$ fog   &                    &          \\
CS7        &   120,124          & 15            & Water  &  Solid 7/8~in slug  &                    &          \\
CS7.2      &   120              & 17            & Water  &  Solid 7/8~in slug  &                    &          \\
CS8        &   120,116          & 15            & Water  &  30$^{\circ}$ fog   &                    &          \\
CS9        &   120,116          & 16            & Water  &  30$^{\circ}$ fog   &                    &          \\
CS10       &   120,116          & 16            & Water  &  Solid Stream       &                    &          \\
CS11       &   120,117          & 16            & Water  &  Solid Stream       &                    &          \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

After weighing each of the bins the mass of water (kg) in each bin was plotted by position to visualize the spray distribution. Figure~\ref{fig:Burn_Building_Test_1} shows that the majority of the

@@

\appendix

\chapter{Fuel Load Details}
\label{app:fuel_loads}
